New approaches to teachers' experience of stress: Do heart rate measurements with fitness trackers provide an efficient, inexpensive, and robust measurement method?

### 2024-06-24 02_paper/02_study/02_fitbit paper/fitbit_paper_elsevier/FitBit_elsevier/FitBit_elsevier.Rmd
tabellen

  <!-- https://tinyurl.com/ybremelq -->
keywords: 
  - heart rate
  - photoplethysmography
  - wearable electronic device
  - teaching
journal: "Computers & Education"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: r-references.bib
linenumbers: false
numbersections: true
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: default
---

```{r analysis-preferences}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

library(needs)
needs(ltm,
      broom,
      ppcor,
      jtools,
      lm.beta,
      tidyverse,
      janitor,
      lubridate,
      readxl,
      ggthemes,
      gridExtra,
      imputeTS,
      DescTools,
      cowplot,
      rstatix,
      ggpubr,
      lme4,
      viridis)
```

```{r data wrangling, include=FALSE}
# disruption & confidence rating
df_rating <- 
  excel_sheets("data/Coding_SRI.xlsx") %>% 
  map_df(~read_xlsx("data/Coding_SRI.xlsx",.)) %>% # read in data with two sheets
  dplyr::select(ID, # select relevant columns 
         disruption_factor, # -99 = subject did not notice event; -100 = experimenter did not asked for event
         confidence_factor,
         event
         ) %>% 
  filter(!ID %in% c("126", # exclude cases with no fitbit data (to check see data Heart Rate)
                    "132",
                    "236")
         ) %>% 
  mutate(disruption_factor = ifelse(disruption_factor < 0,
                                    yes = NA,
                                    no = disruption_factor
                                    ),
         confidence_factor = ifelse(confidence_factor < 0,
                                   yes = NA,
                                   no = confidence_factor),
         event = as_factor(event)
         )

# demographic data
df_demo_all <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% # read in excel 
  transmute(ID = LI06_05, # select and rename relevant columns 
            # gender = factor(LI02_01_1,
            #                 levels = 1:2,
            #                 labels = c("male","female")
            #                 ),
            gender = LI02_01_1, # 1 = male; 2 = female
            age = LI03_01, 
            teaching_experience = LI04_01,
            school_type = LI18,
            term = LI11_01,
            extracurricular_teaching_experience = LI14,
            secondary_teaching_activities = LI17) %>% 
  filter(!ID %in% c("126", # exclude cases with no fitbit data (to check see data Heart Rate)
                    "132",
                    "236")
        )

# merge two data frames by ID 
df_merge <- merge(df_demo_all,
                  df_rating,
                  by = "ID")

@@

data_path <- "./data/heart_rate_data"

main <-
map(.x = dir(path = data_path,
             pattern = ".csv"),
    ~ read_csv(file.path(data_path, .), 
               id = "id",
               col_types = c("t","n")
               ) %>% 
  mutate(time = .$Time - min(.$Time),
         time = as.numeric(time),
         heart_rate = `Heart Rate`,
         ID = id,
         time_span = case_when(str_detect(string = id, pattern = "_a") ~ "end_phase",
                               str_detect(string = id, pattern = "_m") ~ "teaching_phase",
                               str_detect(string = id, pattern = "_p") ~ "pre-teaching_phase",
                               str_detect(string = id, pattern = "_s") ~ "post-teaching_phase",
                               str_detect(string = id, pattern = "_i") ~ "interview_phase",
                               TRUE ~ "overall"
                               ),
         ID = str_extract(string = ID,
                          pattern = "[:digit:]{3}"),
         ID = as.numeric(ID)
      ) %>%
      # filter(time <= 600) %>% # filter for 10min intervals
    filter(time <= 7200) %>% # filter for 2 hours 30min (maximal duration of study)
    dplyr::select(!c("Time","Heart Rate","id"))
    ) %>%
  bind_rows()

@@

main <- 
  main %>% 
  dplyr::select("time", "heart_rate", "ID", "time_span") %>% 
  filter(!ID %in% c("126", # exclude cases with no fitbit data (to check see data Heart Rate)
                    "132",
                    "236")
        ) %>% 
  distinct() %>% 
  group_by(ID) %>% 
  summarize(mean_heart = mean(heart_rate),
            sd_heart = sd(heart_rate)) %>%
  ungroup() %>% 
  right_join(x = .,
             y = main,
             by = "ID") %>% 
  mutate(heart_rate_std = (heart_rate - mean_heart)/
           sd_heart) %>% 
  filter(!(time_span == "teaching" & time > 600))

@@

>>>>

  <!-- https://tinyurl.com/ybremelq -->
keywords: 
  - heart rate
  - photoplethysmography
  - wearable electronic device
  - teaching
journal: "Computers & Education"
date: "`r Sys.Date()`"
classoption: preprint, 3p, authoryear
bibliography: r-references.bib
linenumbers: false
numbersections: true
output: 
  rticles::elsevier_article:
    keep_tex: true
    citation_package: natbib
header-includes:
  - \usepackage{tabularx}
---

```{r analysis-preferences, include=FALSE}
# Seed for random number generation
set.seed(42)
knitr::opts_chunk$set(cache.extra = knitr::rand_seed)
```

@@

library(needs)
needs(ltm,
      broom,
      ppcor,
      jtools,
      lm.beta,
      tidyverse,
      janitor,
      lubridate,
      readxl,
      ggthemes,
      gridExtra,
      imputeTS,
      DescTools,
      cowplot,
      rstatix,
      ggpubr,
      lme4,
      viridis, 
      stargazer)
```

```{r data wrangling, include=FALSE}
# disruption & confidence rating

df_rating <- 
  excel_sheets("data/Coding_SRI.xlsx") %>% 
  map_df(~ read_xlsx("data/Coding_SRI.xlsx",.)) %>% # read in data with two sheets
  dplyr::select(ID, # select relevant columns 
         "disruption_factor", # -99 = subject did not notice event; -100 = experimenter did not asked for event
         "confidence_factor",
         "event"
         ) %>% 
  mutate(disruption_factor = ifelse(disruption_factor < 0,
                                    yes = NA,
                                    no = disruption_factor
                                    ),
         confidence_factor = ifelse(confidence_factor < 0,
                                   yes = NA,
                                   no = confidence_factor),
         event = as_factor(event)
         )

# demographic data

df_demo <- 
  read_excel("./data/data_empschul_labor_lehrperson.xlsx") %>% # read in excel 
  transmute(ID = LI06_05, # select and rename relevant columns 
            gender = factor(LI02_01_1,
                            levels = 1:2,
                            labels = c("male","female")
                            ),
            age = LI03_01, # 1 = male; 2 = female
            teaching_experience = LI04_01)

# merge two data frames by ID

df_merge <- merge(df_demo,
                  df_rating,
                  by = "ID") %>% 
  filter(!ID %in% c("126", # exclude cases with no fitbit data (to check see data Heart Rate)
                    "132")
         )

@@

data_path <- "data/heart_rate_data"

main <-
  map(.x = dir(path = data_path,
               pattern = ".csv"),
      ~ read_csv(file.path(data_path, .), 
                 id = "id",
                 col_types = c("t","n")
                 ) %>% 
  mutate(time = .$Time - min(.$Time),
         time = as.numeric(time),
         heart_rate = `Heart Rate`,
         ID = id,
         time_span = case_when(str_detect(string = id, pattern = "_a") ~ "end",
                               str_detect(string = id, pattern = "_m") ~ "teaching",
                               str_detect(string = id, pattern = "_p") ~ "preparation",
                               str_detect(string = id, pattern = "_s") ~ "post",
                               str_detect(string = id, pattern = "_i") ~ "interview",
                               TRUE ~ "overall"
                               ),
         ID = str_extract(string = ID,
                          pattern = "[:digit:]{3}"),
         ID = as.numeric(ID)
      ) %>%
    # filter(time <= 600) %>% # filter for 10min intervals
    filter(time <= 7200) %>% # filter for 2 hours (maximal duration of study)
    dplyr::select(!c("Time","Heart Rate","id"))
    ) %>%
  bind_rows()

@@

main <- 
  main %>% 
  dplyr::select("time", "heart_rate", "ID", "time_span") %>% 
  distinct() %>% 
  group_by(ID) %>% 
  dplyr::summarise(mean_heart = mean(heart_rate),
            sd_heart = sd(heart_rate)) %>%
  ungroup() %>% 
  right_join(x = .,
             y = main,
             by = "ID") %>% 
  filter(!(time_span == "teaching" & time > 600)) %>%
  mutate(heart_rate_std = (heart_rate - mean_heart)/
           sd_heart)
```

```{r descriptives, include=FALSE}
# Mean, SD and range for overall (unstandardized)

descrip_hr_overall <- main %>%
  filter(time_span %in% c("overall")) %>% 
  summarise(N = n_distinct(ID),
            "M heart rate in bpm" = round(mean(heart_rate),
                                     digits = 2),
            "SD heart rate in bpm" = round(sd(heart_rate),
                                      digits = 2),
            "Min heart rate in bpm" = min(heart_rate),
            "Max heart rate in bpm" = max(heart_rate)
  )

# Mean, SD and range for overall (standardized)

descrip_hr_overall_std <- main %>%
  filter(time_span %in% c("overall")) %>% 
  summarise(N = n_distinct(ID),
            "M std heart rate in bpm" = round(mean(heart_rate_std),
                                              digits = 2),
            "SD std heart rate in bpm" = round(sd(heart_rate_std),
                                               digits = 2),
            "Min std heart rate in bpm" = round(min(heart_rate_std),
                                                digits = 2),
            "Max heart rate in bpm" = round(max(heart_rate_std),
                                            digits = 2)
            )

# Mean, SD and range for individual phases (unstandardized)

descrip_hr_phases <- main %>% 
  filter(!time_span %in% c("overall")) %>% 
  group_by(time_span) %>% 
  summarise(N = n_distinct(ID),
            "M heart rate in bpm" = round(mean(heart_rate),
                                     digits = 2),
            "SD heart rate in bpm" = round(sd(heart_rate),
                                      digits = 2),
            "Min heart rate in bpm" = min(heart_rate),
            "Max heart rate in bpm" = max(heart_rate)
  )

# Mean, SD and range for individual phases (standardized)

descrip_hr_phases_std <- main %>% 
  filter(!time_span %in% c("overall")) %>% 
  group_by(time_span) %>% 
  summarise(N = n_distinct(ID),
            "M heart rate in bpm" = round(mean(heart_rate_std),
                                     digits = 2),
            "SD heart rate in bpm" = round(sd(heart_rate_std),
                                      digits = 2),
            "Min heart rate in bpm" = round(min(heart_rate_std),
                                            digits = 2),
            "Max heart rate in bpm" = round(max(heart_rate_std),
                                            digits = 2)
  )

# # format and insert table in manuscript
# descrip_hr_phases %>%
# papaja::apa_table(
#   caption = "Mean, SD and range for teachers' heart rate in individual phases",
#   # note = "Write Note here",
#   escape = TRUE, # if TRUE special Latex characters are escaped; if this is turned to F captions cannot be rendered. I don't know why...
#   placement = "h", # position of table in page:  exact location (h), at the top (t), bottom (b)
#   font_size = "tiny" # options are tiny, scriptsize, footnotesize, small, normalsize (default), large, Large, LARGE, huge, Huge
#   )

# Mean, SD and range for disruption and confidence factor

rating_table <- 
  df_rating %>% 
  filter(!is.na(confidence_factor),
         !is.na(disruption_factor)) %>% 
   summarise(N = n_distinct(ID), 
            "M disruption factor" = round(mean(disruption_factor),
                                     digits = 2),
            "SD disruption factor" = round(sd(disruption_factor),
                                      digits = 2),
            "Min disruption factor" = min(disruption_factor),
            "Max disruption factor" = max(disruption_factor),
            "M confidence factor" = round(mean(confidence_factor),
                                     digits = 2),
            "SD confidence factor" = round(sd(confidence_factor),
                                      digits = 2),
            "Min confidence factor" = min(confidence_factor),
            "Max confidence factor" = max(confidence_factor)
            )
```

```{r, include=FALSE}

desc_hr_phases <- bind_rows(x = descrip_hr_overall,
                            y = descrip_hr_phases) |> 
  select(-N) |> 
  rename(`M HR` = `M heart rate in bpm`,
         `SD HR` = `SD heart rate in bpm`,
         `Min` = `Min heart rate in bpm`,
         `Max` = `Max heart rate in bpm`,
         Interval = time_span) |> 
  select(c(Interval,
           `M HR`,
           `SD HR`,
           `Min`,
           `Max`)
         )

descrip_hr_overall_std <- descrip_hr_overall_std |> 
  rename(`M HR` = `M std heart rate in bpm`,
         `SD HR` = `SD std heart rate in bpm`,
         `Min` = `Min std heart rate in bpm`,
         `Max` = `Max heart rate in bpm`,
          ) |> 
  mutate(Interval = 'overall')

descrip_hr_phases_std <-  descrip_hr_phases_std |> 
  rename(`M HR` = `M heart rate in bpm`,
         `SD HR` = `SD heart rate in bpm`,
         `Min` = `Min heart rate in bpm`,
         `Max` = `Max heart rate in bpm`,
         Interval = time_span)

desc_hr_phases_std <- bind_rows(x = descrip_hr_overall_std,
                                y = descrip_hr_phases_std) |> 
  select(-N) |> 
    select(c(Interval,
           `M HR`,
           `SD HR`,
           `Min`,
           `Max`)
         )

desc_hr_phases[1, 1] <- 'overall'

df_desc_hr <- desc_hr_phases |> 
  mutate(across(everything(),
                ~ paste(.,
                        desc_hr_phases_std[[cur_column()]],
                        sep = "/")
                )
         ) #|> 
  # # rename(`\textit{M} HR` = `M HR`,
  #        # `\textit{SD} HR` = `SD HR`,
  #        # `\textit{Min}` = `Min`,
  #        `\textit{Max}` = `Max`,
  #       `\textit{Interval}` = Interval)

df_desc_hr$Interval <- c('Overall Course of 2h',
                         'Pre-teaching interval',
                         'Teaching interval',
                         'Post-teaching interval',
                         'Interview interval',
                         'End Interval')

names(df_desc_hr) <- paste0("\\textit{", names(df_desc_hr),"}")

stargazer(df_desc_hr,
          type = 'latex',
          out = 'tab_1.tex',
          summary = FALSE,
          rownames = FALSE,
          table.placement = 'h',
          align = FALSE
          ) -> tab_1

pattern <- "textit\\\\\\{([^}]*)\\\\\\}"
replacement <- "\textit{\\1}"

# Anwenden des Regex
tab_1_new <- str_replace_all(tab_1, pattern, replacement)

paste(tab_1_new, collapse = '')

writeLines(tab_1_new, 'tab_1.tex')

@@

```{=latex}
\input{tab_1.tex}
```

### 2024-06-24 02_paper/02_study/02_fitbit paper/fitbit_paper_elsevier/FitBit_elsevier/FitBit_elsevier.tex
tabellen

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}

@@

% Pandoc citation processing
\newlength{\cslhangindent}
\setlength{\cslhangindent}{1.5em}
\newlength{\csllabelwidth}
\setlength{\csllabelwidth}{3em}
\newlength{\cslentryspacingunit} % times entry-spacing
\setlength{\cslentryspacingunit}{\parskip}
% for Pandoc 2.8 to 2.10.1
\newenvironment{cslreferences}%
  {}%
  {\par}
% For Pandoc 2.11+
\newenvironment{CSLReferences}[2] % #1 hanging-ident, #2 entry spacing
 {% don't indent paragraphs
  \setlength{\parindent}{0pt}
  % turn on hanging indent if param 1 is 1
  \ifodd #1
  \let\oldpar\par
  \def\par{\hangindent=\cslhangindent\oldpar}
  \fi
  % set entry spacing
  \setlength{\parskip}{#2\cslentryspacingunit}
 }%
 {}
\usepackage{calc}
\newcommand{\CSLBlock}[1]{#1\hfill\break}
\newcommand{\CSLLeftMargin}[1]{\parbox[t]{\csllabelwidth}{#1}}
\newcommand{\CSLRightInline}[1]{\parbox[t]{\linewidth - \csllabelwidth}{#1}\break}
\newcommand{\CSLIndent}[1]{\hspace{\cslhangindent}#1}

@@

\hypertarget{present-investigation}{%
\section{Present investigation}\label{present-investigation}}

@@

According to previous findings that fitness trackers can be used as a
low-cost, non-invasive method of measuring HR (\textbf{hajj2022wrist?};
Fuller et al. 2020) and that different HRs of teachers can be measured
in different teaching phases (\textbf{donker2020associations?};
\textbf{junker2021potential?}), we formulate the hypotheses as follows:

@@

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\vadjust pre{\hypertarget{ref-fuller2020reliability}{}}%
Fuller, Daniel, Emily Colwell, Jonathan Low, Kassia Orychock, Melissa
Ann Tobin, Bo Simango, Richard Buote, et al. 2020. {``Reliability and
Validity of Commercially Available Wearable Devices for Measuring Steps,
Energy Expenditure, and Heart Rate: Systematic Review.''} \emph{JMIR
mHealth and uHealth} 8 (9): e18694.

\end{CSLReferences}

>>>>

\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage{amssymb,amsmath}
% TODO: Currently lineno needs to be loaded after amsmath because of conflict
% https://github.com/latex-lineno/lineno/issues/5
\usepackage{lineno} % add
\usepackage{ifxetex,ifluatex}
\usepackage{fixltx2e} % provides \textsubscript
% use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\ifnum 0\ifxetex 1\fi\ifluatex 1\fi=0 % if pdftex
  \usepackage[utf8]{inputenc}
\else % if luatex or xelatex
  \usepackage{fontspec}
  \ifxetex
    \usepackage{xltxtra,xunicode}
  \fi
  \defaultfontfeatures{Mapping=tex-text,Scale=MatchLowercase}
  \newcommand{\euro}{€}
\fi
% use microtype if available
\IfFileExists{microtype.sty}{\usepackage{microtype}}{}
\usepackage[]{natbib}
\bibliographystyle{plainnat}

@@

\usepackage{tabularx}

@@

% Table created by stargazer v.5.2.3 by Marek Hlavac, Social Policy Institute. E-mail: marek.hlavac at gmail.com
% Date and time: Mo, Jun 24, 2024 - 16:22:09

\\begin{table}[h] 
\\centering 
\\caption{} 
\\label{}
\\begin{tabular}{@{
\\extracolsep{5pt}} ccccc} \\\\[-1.8ex]\\hline \\hline \\\\[-1.8ex] \\textbackslash \textit{Interval} & \\textbackslash \textit{M HR} & \\textbackslash \textit{SD HR} & \\textbackslash \textit{Min} & \\textbackslash \textit{Max} \\\\ \\hline \\\\[-1.8ex] Overall Course of 2h & 90.09/-0.04 & 15.76/0.99 & 51/-4.03 & 164/4.56 \\\\ Pre-teaching interval & 77.95/-1.07 & 11.14/0.57 & 50/-2.68 & 120/2.96 \\\\ Teaching interval & 82.32/-0.72 & 11.85/0.74 & 51/-2.51 & 132/4.39 \\\\ Post-teaching interval & 93.61/0.27 & 14.01/0.76 & 60/-2.17 & 150/3.06 \\\\ Interview interval & 96.28/0.48 & 14.11/0.88 & 56/-3.56 & 139/3.24 \\\\ End Interval & 100.8/0.85 & 16.23/0.77 & 63/-2.18 & 164/4.37 \\\\ \\hline \\\\[-1.8ex]
\\end{tabular}
\\end{table}

%\input{tab_1.tex}

\section{Present investigation}\label{present-investigation}

@@

According to previous findings that fitness trackers can be used as a
low-cost, non-invasive method of measuring HR
\citep{hajj2022wrist, fuller2020reliability} and that different HRs of
teachers can be measured in different teaching phases
\citep{donker2020associations, junker2021potential}, we formulate the
hypotheses as follows:

@@

\renewcommand\refname{References}
\bibliography{r-references.bib}
